What's hot and what's not:\\Windowed developer topic analysis

### 2009-06-26 lda-paper.tex
migod edits

\begin{abstract}

  As development on a software project progresses, developers shift
  their focus between different topics and tasks many times. Managers
  and newcomer developers often seek ways of understanding what tasks
  have recently been worked on and how much effort has gone into each;
  for example, a manager might wonder what unexpected tasks occupied
  their team's attention during a period when they were supposed to
  have been implementing a set of new features.  Tools such as Latent
  Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) can be
  used to extract a set of independent topics from a corpus of commit
  log comments.  Previous work in the area has created a single set of
  topics by analyzing comments from the entire lifetime of the
  project.  In this paper, we propose windowing the topic analysis to
  give a more nuanced view of the system's evolution. By using a
  defined time-window of, for example, one month, we can track which
  topics come and go over time, and which ones recur. We propose
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time. We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a project's lifetime because many topics are quite
  local.

@@

  The questions that we wish to address deal with the past and
  topics development that occurred in the past.  What happened in
  the previous iteration of a project's development?  What were the
  developers working on? What were the topics of the
  previous release? Given the requirements agreed to at the start of
  the iteration did our developers work on them?  What else did they
  work on?

We want to know about the topics of that programmers
encountered while developing a software project. We want to know given
a certain period of time, what were the popular topics being worked on
during that period. We also want to know if any of those topics
occurred again, whether before or after our current period.  These
topics could help stakeholders determine what their coworkers were
working on, what were common issues, what requirements were being
worked on, etc. In order to extract and evaluate these topics we plan
to rely on topic analysis.

Topic analysis extracts independent word distributions (topics) from
documents (commit log comments).
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. 
Topic analysis often allows for a single document, such as a commit
message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS).

In the past, topic analysis has been applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
where-as this paper suggests that most topics are local. For example,  
programmer may be working on a story card, they might never deal with that
story card ever again. Thus that topic might be quite relevant to a
small window of time but never relevant ever again.  By applying topic
analysis locally to a window or an iteration, both the developers and managers
might be able to see what were the topics of development during that
iteration, rather than the whole project.

We want to see topics that are unique to a
certain time period, displayed across the time axis.
Figure \ref{fig:lda} illustrates the kind of visualization we want to
automatically generate. Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a word drawn from its
word distribution.  
Given a visualization, such as Figure \ref{fig:lda}, topic
analysis can aid in partitioning a project's time-line into periods of
developer focus.
By breaking apart an iteration into sets of topics
and trends (topics that recur), we may be able to recognize the
underlying software development process and maintenance tasks from these
commits.

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by developers. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e. a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing
%(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
(LSI)~\cite{1374321,10.1109/ICPC.2007.13},
probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
semantic clustering~\cite{1698774,1566153}. These tools are used for
document modelling, document clustering and collaborative
filtering. LDA attempts to encode documents as a mixture
model, a combination of topics.  LDA is used in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20}%,NIPS2007637,1321709}.
LDA extracts topics from documents such as methods, bug
reports, source code and files.

LDA, LSI and semantic clustering extract topic
clusters from the documents that are independent from one another. LDA
does not name these clusters but the clusters consist of words whose
relationship is often obvious to someone familiar with the corpus. For our purposes, we
could swap LDA for LSI or semantic clustering and likely produce similar
results. Our data is posed as documents with word distributions (word
counts per documents) and LDA extracts distinct topics (clusters of
words) from the documents.

@@

Our technique differs from the previous techniques because we apply
LDA locally to month-long windows of commit log comments, whereas
other approaches apply LDA once to the entire project. Windowed topic
analysis allows us to examine the time-based windowing of topics over
its development history.

@@

Imagine that we are analyzing one month of development in a project
where the main focus was on bug fixing, adding a new GUI and
documentation updates relating to the system and the GUI. To use LDA
we would get the word distributions of each commit messages, pass them to
LDA and tell LDA to find some topics, in this case we
in this case we chose to concentrate on four topics.
 The four
topics it discovered would ideally be bug fixing, GUI, documentation
and other. These topics are essentially word distributions, for
example the bug fixing topic would include words like \emph{bug},
\emph{fix}, \emph{failure} and \emph{ticket}. The GUI topic would
include words like \emph{widget}, \emph{panel}, \emph{window}, and
\emph{menu}. The documentation topic would include words like
\emph{section}, \emph{figure}, and \emph{chapter}. The other topic
would include words found in topics generally independent of the other
topics. If we analyzed a commit of a bug fix it would probably contain
words like \emph{ticket} and thus heavily associate with the bug fix
topic. A commit that was about user documentation regarding the GUI
would be associated with the documentation topic and the GUI topic
because it shared words from both topics. Where as a topic dealing
neither bugs, GUIs or documentation might be associated with the other
topic. This example demonstrates how LDA can associate one document (a
commit message) with more than one topic, and how topics are related to
documents via their word distributions. LDA also would discover these
independent topics in an unsupervised manner. All the end user has to
do is name the topic based upon the most frequent words in that topic,
which is a word distribution.

@@

We found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}; since these words
occur so often in the MySQL 3.23 messages. These words probably should
be viewed as stop words since occur across so many topics, even though
they are related to version control usage.  There were notable
transitional topics such as in the first window the word
\emph{BitKeeper} appears because MySQL adopted Bitkeeper for their
source control system yet in the following windows there were no
mentions of BitKeeper. \emph{RENAME} also only appeared once in the
first window and never again. For each topic we tried to name the
purpose of the topic; to our surprise we found that even with only a
little familiarity with the code base that naming the topic was
straightforward. To find the purpose of a commit we looked at the most
frequent words in the word distribution of the topic and tried to
summarize the topic, then we looked into the commits related to that
topic to investigate if we were correct; since the commit messages and
the word distribution share the same words the purpose extracted from
commit comments was usually accurate.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. This is the
    kind of plot we eventually want to illustrate: named topics and
    topic trends. The horizontal axis is time by month examined. The
    vertical axis is used to stack topics that occur at the same
    time. Longer topics are topics which recur in adjacent
    windows. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into topics and trends. Commits are first extracted, then abstracted into word counts or word distributions. These word distributions are then given to topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to (the numbers indicate similarity between documents and topics).}
  \label{fig:commits}
\end{figure*}

@@

The documents we are analyzing are the commit log comments, i.e. the
comments that are added when revisions to the project are committed.
For each commit log comment, we count the occurrence of each word in
the message, remove stop words, and then produce word distributions
for each message.  These distributions are then normalized by the size
of the message, each count was divided by the total number of
tokens. After all messages are processed the distributions are
extended to include all words from all of the distributions.

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics. We used $20$ topics in our
case studies because we experimented before and that $20$ was about
all we can handle before too many topics seemed too indistinct from
each other to differentiate.

 Our Topic Analysis tool is an implementation of LDA.
We could have used other similar tools like LSI, but previous work
showed more promising topic analysis results with LDA.  We note that
this step is a slow one, as executing even one instance of LDA
involves significant computation, and we perform LDA once per window.
Figure \ref{fig:commits} shows how LDA is applied to a set of
messages, and how the topics are extracted and related to the
messages.

@@

Our input for topic similarity is the top 10 most common words of each
topic.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 word similarity, such as 8 out of
top 10 matching words.  10 words were chosen because often people care
about top 10 comparison and it allowed for some common words to exist
but not cause topics to match so closely.  We then apply the
transitive closure on similar topics to this topic similarity matrix;
this is similar to modelling topics as nodes and similarity as arcs,
then fill flooding along the similarity arcs until we have partitioned
the topics into clusters of similar topics. Figure \ref{fig:closure}
illustrates clustering of topics by topic similarity. These clusters
of topics are called trends. 
Trends are interesting because they that topic during development has occurred over more than one period.

This technique has weaknesses in that nodes that are a few neighbors
away in similarity might not share any similar words.  We use this
measure because we want to be able to color or highlight topics that
are related to each other and track them throughout time.

Once we have determined our similarity clusters we are ready analyze
and to plot the topics.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.42\textwidth]{transitiveclosure}
  \caption{Topic similarity demonstrated by clustering topics by the transitive closure (connectedness)
    of topic similarity. Nodes are topics and
    arcs imply similarity. Similarity could be a measure such as
    topics share 8 out of 10 top ten words.}
\label{fig:closure}
\end{figure}

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer
  Visualization is an integral part of our topics analysis which
  allows us to quickly explore the topics and trends of a project.
  Visualization provides us with a framework that allows us to
  visually answer many questions about: the spread of topics, how many
  topics were independent per period, what the repeating trends were,
  if the trends dominated or did local topics dominate, how continuous
  trends were, and what the topic text in topic, trend, or period was.
  Since we have multiple questions to answer we have multiple
  visualization, described within this section, that help answer
  different questions.

We have devised several techniques for visualizing these topics and
trends.  For all of these techniques if we find trends that have
continuous segments, then we plot those segments as joined
horizontally across time. One technique is the \emph{compact-trend-view},
shown in Figure \ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear},
that displays trends as they occur on the time-based x-axis
(placement along the y-axis is arbitrary).  Another technique is
the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline},
each trend gets it own distinct y-value, while the x-axis is time;
these topics are then each plotted on their own line across time as
they occur. Our final technique is the \emph{trend-histogram}, shown in Figure
\ref{fig:histogram} where we plot each trend on its own line but stack
up the segments of the trend, much like a histogram. Each topic has its top 10 words 
listed in descending order of frequency from top to bottom, this text is embedded inside the topic's box.
A trend has all of its topic text embedded side by side within the same trend box.

 The \emph{compact-trend-view} (Figure \ref{fig:topicsmear})
  attempts to show all topics and trends at once across time in a
  compact view that could fit on one page. It tries to give a compact
  summary of the topics and trends across time.   The
\emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to sink
the larger trends to the bottom.  Once the larger trends are stacked
we fill in the gaps with the smaller trends in the same window, and
then stack the smaller trends on top.  Although there is a chance that
gaps will be left due to non-optimal stacking, in practice there are
lot of small trends (90\% of all trends contain 1 topic) that fill in
these gaps quickly.  Different instances of the same trend share the
same color; apart from that, the color of trends is randomly chosen
and color similarity is not meaningful.  The compact-trend-view is
convenient because it shows all of the topic information, as shown by
the zoomed-in view of MaxDB 7.500's compact-trend-view in Figure
\ref{fig:zoomedsmear}. The \emph{compact-trend-view} makes repeating
continuous trends easy to pick out, although discontinuous trends are
harder to spot, and provides a general summary of the topics within a project.

@@

We applied our tools and visualizations to the repositories of these
open source database systems: PostgreSQL, MaxDB and Firebird. The
total number of commits analyzed was over $66 000$.

@@

When we examined PostgreSQL's history from 1996 to 2004, which includes
over $20 000$ commits. We did not find many trends with two or more
topics, using a similarity of $7/10$. $7/10$ was chosen because it
preserve independent topics but seemed to be a threshold
value where more serious trends started to appear.

@@

We tracked Firebird from August 2000 to January 2006, we extracted comments from $38 000$ commits. We found that with a
similarity of $7/10$ Firebird had far more continuous and
recurring trends than PostgreSQL.  The first large trend was discontinuous
across time but explicitly references one author \texttt{carlosga05}
and words like \emph{added}, \emph{fixed} and \emph{updated}.

@@

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur
four times, yet in the global-trend-view (Figure
\ref{fig:statictopics}, explained in Section \ref{sec:compare}) it is not in any of the topics. A query
optimizer is an important component of an DBMS, but as we have shown
it does not appear as a topic on its own. We tried to remove words to
see if we could get an \emph{optimizer} topic. After removing stop
words and then two of the most common words, the global analysis
finally had a topic with optimizer in its topic ten words. Our
analysis shows that optimizer was important but it had been obscured
by the global topic analysis, which used the entire history of
messages, but would have been noticed using the more local topic
analysis such as our windowed topic analysis, which used a window of
messages.

>>>>

%XXX Drop down to 200
\begin{abstract}

  As development on a software project progresses, developers shift
  their focus between different topics and tasks many times. Managers
  and newcomer developers often seek ways of understanding what tasks
  have recently been worked on and how much effort has gone into each;
  for example, a manager might wonder what unexpected tasks occupied
  their team's attention during a period when they were supposed to
  have been implementing new features.  
%  have been implementing a set of new features.  
%
  Tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic
%  Indexing (LSI) can be used to extract a set of independent topics
  Indexing (LSI) extract a set of independent topics
  from a corpus of commit-log comments.  Previous work in the area has
  created a single set of topics by analyzing comments from the entire
  lifetime of the project.  In this paper, we propose windowing the
  topic analysis to give a more nuanced view of the system's
  evolution. By using a defined time-window of, for example, one
  month, we can track which topics come and go over time, and which
  ones recur.
%
%
%   Tools such as Latent
%   Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) can be
%   used to extract a set of independent topics from a corpus of commit
%   log comments.  Previous work in the area has created a single set of
%   topics by analyzing comments from the entire lifetime of the
%   project.  In this paper, we propose windowing the topic analysis to
%   give a more nuanced view of the system's evolution. By using a
%   defined time-window of, for example, one month, we can track which
%   topics come and go over time, and which ones recur.
%  
  We propose
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time. We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a project's lifetime because many topics are quite
  local.

@@

Software managers know that --- during any given development cycle ---
there is usually a difference between the set of tasks that developers
are supposed to be working on and the tasks they are actually working
on.  To better understand the realities of development, it would be
helpful to be able to answer questions such as: What topics dominated
the previous release cycle? Given the requirements agreed to at the
start of the iteration how much work did our developers spend on them?
What other tasks did they work on?  Which topics seem to recur over
and over between cycles?  How long do different topics persist?  Are
there latent cause-effect relationships between different topics?

%   The questions that we wish to address deal with the past and
%   topics development that occurred in the past.  What happened in
%   the previous iteration of a project's development?  What were the
%   developers working on? What were the topics of the
%   previous release? Given the requirements agreed to at the start of
%   the iteration did our developers work on them?  What else did they
%   work on?

% We want to know about the topics of that programmers
% encountered while developing a software project. We want to know given
% a certain period of time, what were the popular topics being worked on
% during that period. We also want to know if any of those topics
% occurred again, whether before or after our current period.  These
% topics could help stakeholders determine what their coworkers were
% working on, what were common issues, what requirements were being
% worked on, etc. In order to extract and evaluate these topics we plan
% to rely on topic analysis.

Topic analysis 
uses tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
documents (commit log comments)~\cite{1374321,10.1109/ICPC.2007.13,lukins2008,10.1109/MSR.2007.20}.
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. 
Topic analysis often allows for a single document, such as a commit
message, to be related to multiple topics. Documents represented as a mixture of topics maps well to commits
to source code, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this case a topic
is a set of tokens extracted from commit messages found within a
projects source control system (SCS).

Previous work on topic analysis has been applied to the entire history of a 
project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13}
to produce a single set of topics. In this paper we explore the idea
of time-windowed topic analysis; that is, we pick a period of, say,
one month and perform topic analysis within each window.  In addition
to being able to study just which topics were actually worked on and
when, we can also analyze the historical timeline of topics for
patterns and trends.  For example, we might wish to categorize each
topic as being local to a single time window, spanning multiple
consecutive windows, or recurring periodically.  Alternatively, we
might explore if any topics often recur together, if one particular
topic usually precedes another, if unplanned topics outnumber planned
topics, if particular topics recur near the end of development
iterations, etc.  whereas this paper suggests that most topics are
local.

% In the past, topic analysis has been applied to the entire history of a project~\cite{1321709,lukins2008,10.1109/ICPC.2007.13},
% whereas this paper suggests that most topics are local. For example,  
% programmer may be working on a story card, they might never deal with that
% story card ever again. Thus that topic might be quite relevant to a
% small window of time but never relevant ever again.  By applying topic
% analysis locally to a window or an iteration, both the developers and managers
% might be able to see what were the topics of development during that
% iteration, rather than the whole project.

% We want to see topics that are unique to a
% certain time period, displayed across the time axis.
% Figure \ref{fig:lda} illustrates the kind of visualization we want to
% automatically generate. 
As part of this work, we wish to explore approaches to automatically
generating visualizations of the topic trends and development timelines, so
that managers may be able to more easily grasp the topic-related
activities of the developers.  Figure \ref{fig:lda} gives an example.
Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a word drawn from its
word distribution.  
Given a visualization, such as Figure \ref{fig:lda}, topic
analysis can aid in partitioning a project's time-line into periods of
developer focus.
By breaking apart an iteration into sets of topics
and trends (topics that recur), we may be able to recognize the
underlying software development process and maintenance tasks from these
commits.

@@

We now define some terms that we will use in the rest of the paper: A
\emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

@@

In terms of clustering and finding topic distributions, Latent
Dirichlet Allocation (LDA)~\cite{944937} competes with Latent Semantic
Indexing
%(LSI)~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17},
(LSI)~\cite{1374321,10.1109/ICPC.2007.13},
probabilistic Latent Semantic Indexing (pLSI)~\cite{944937} and
semantic clustering~\cite{1698774,1566153}. These tools are used for
document modelling, document clustering and collaborative
filtering. LDA attempts to encode documents as a mixture
model, a combination of topics.  LDA has been used in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20}%,NIPS2007637,1321709}.
to extract topics from documents such as methods, bug
reports, source code and files.

LDA, LSI and semantic clustering extract topic
clusters from the documents that are independent from one another. 
The clusters are not named by the process, and consist of
%LDA
%does not name these clusters but the clusters consist of
words whose
relationship is often obvious to someone familiar with the corpus. For our purposes, we
could swap LDA for LSI or semantic clustering and likely produce similar
results. Our data is posed as documents with word distributions (word
counts per documents) and LDA extracts distinct topics (clusters of
words) from the documents.

@@

Our technique differs from the previous approaches to topic analysis because we apply
LDA locally to month-long windows of commit log comments, whereas
other approaches apply LDA once to the entire project. Windowed topic
analysis allows us to examine the time-based windowing of topics over
its development history.

@@

Imagine that we are analyzing one month of development in a project
where there were three major tasks: bug fixing, adding a new GUI, and
documentation updates relating to the system and the GUI. To use LDA
we would get the word distributions of each commit messages, 
\begin{comment}
pass them to
LDA and tell LDA to find some topics, in this case we
in this case we chose to concentrate on four topics.
 The four
topics it discovered would ideally be bug fixing, GUI, documentation
and other. These topics are essentially word distributions, for
example the bug fixing topic would include words like \emph{bug},
\emph{fix}, \emph{failure} and \emph{ticket}. The GUI topic would
include words like \emph{widget}, \emph{panel}, \emph{window}, and
\emph{menu}. The documentation topic would include words like
\emph{section}, \emph{figure}, and \emph{chapter}. The other topic
would include words found in topics generally independent of the other
topics. If we analyzed a commit of a bug fix it would probably contain
words like \emph{ticket} and thus heavily associate with the bug fix
topic. A commit that was about user documentation regarding the GUI
would be associated with the documentation topic and the GUI topic
because it shared words from both topics. Where as a topic dealing
neither bugs, GUIs or documentation might be associated with the other
topic. This example demonstrates how LDA can associate one document (a
commit message) with more than one topic, and how topics are related to
documents via their word distributions. LDA also would discover these
independent topics in an unsupervised manner. All the end user has to
do is name the topic based upon the most frequent words in that topic,
which is a word distribution.
\end{comment}
% (please preserve the italics you put in ... I had to edit this in plain text so they got lost)
and have the LDA tool extract a set of topics from the data.  Ideally,
the LDA analysis would return four topics, corresponding to bug
fixing, GUI implementation, documentation, and ``other''.  Each topic
essentially corresponds to a word distribution.  For example, the bug
fixing topic might include words such as \emph{bug}, \emph{fix}, \emph{failure}, and
\emph{ticket}, while the GUI topic might include \emph{widget}, \emph{panel}, \emph{window}, and
\emph{menu}, and the documentation topic might include \emph{section}, \emph{figure}, and
\emph{chapter}.  The last topic would include words that were not commonly
found in bug reports, GUI fixing, or documentation commits; that is,
the LDA analysis would put into this last category the words that were
independent of the other topics it found.

LDA is an unsupervised technique.  The user provides a corpus of
documents and specifies various parameters, and then LDA automatically
generates a set of likely topics based on word frequencies.  It is
then up to the user to decide if the topics are meaningful by looking
at the most frequent words in each and deciding what each discovered
topic should be named.

In addition to finding the topics, LDA can also associate the
documents (i.e., commit messages) in its corpus to one or more of the
topics it discovers.  For example, a commit of a simple bug fix would
probably contain a word such as \emph{ticket}, and thus heavily associate
with the bug fix topic.  A commit that concerned user documentation of
the GUI would likely be associated with both the documentation and GUI
topics if it shared words from both topics.  And a commit that did not
concern bugs, GUIs, or documentation would likely be related to the other topic.

@@

%   We decided [did we?] to treat these as stop words, since they occur
%   across many of the topics and since they relate to the mechanisms of
%   version control rather than the semantic content of the commit.

We found there were common words across topic clusters, such as
\emph{diffs}, \emph{annotate} and \emph{history}.
Words like these could be treated as stop words, since they occur
across many of the topics and since they relate to the mechanisms of
version control rather than the semantic content of the commit.
% since these words
%occur so often in the MySQL 3.23 messages. These words probably should
%be viewed as stop words since occur across so many topics, even though
%they are related to version control usage.  
There were topics which appeared only during some transitional period and never again
such as \emph{RENAME} and \emph{BitKeeper}. \emph{BitKeeper} which appears when MySQL switched to  BitKeeper for their SCS.
% There were notable
% transitional topics such as in the first window the word
% \emph{BitKeeper} appears because MySQL adopted Bitkeeper for their
% source control system yet in the following windows there were no
% mentions of BitKeeper. 
%\emph{RENAME} also only appeared once in the
%first window and never again.
For each topic we tried to find a word to describe its purpose.
To our surprise we found that even with only a
little familiarity with the code base that naming the topic was
straightforward. To find the purpose of a commit we looked at the most
frequent words in the word distribution of the topic and tried to
summarize the topic, then we looked into the commits related to that
topic to investigate if we were correct; since the commit messages and
the word distribution share the same words the purpose extracted from
commit comments was usually accurate.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{lda}
  \caption{Example of topics extracted from MySQL 3.23. 
%     This is the
%     kind of plot we eventually want to illustrate: named topics and
%     topic trends. 
    The horizontal axis is time by month examined. The
    vertical axis is used to stack topics that occur at the same
    time. Longer topics are topics which recur in adjacent
    windows. Colors are arbitrary.}
  \label{fig:lda}
\end{figure*}

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{commit-to-topics} 
  \caption{How commits are analyzed and aggregated into topics and trends: commits are first extracted, then abstracted into word counts or word distributions which are then given to topic analysis tool like LDA. LDA finds independent word distributions (topics) that these documents are related to (the numbers indicate similarity between documents and topics).}
  \label{fig:commits}
\end{figure*}

@@

The documents we are analyzing are the commit log comments, i.e., the
comments that are added when revisions to the project are committed.
For each commit log comment, we count the occurrence of each word in
the message, remove stop words, and then produce word distributions
for each message.  These distributions are then normalized by the size
of the message, each count was divided by the total number of
tokens. After all messages are processed the distributions are
extended to include all words from all of the distributions.

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics.  We decided to use a
threshold of $20$ topics based on our previous experience with LDA; we
found that using more than $20$ makes the boundaries between topics less
distinct.

% We used $20$ topics in our
% case studies because we experimented before and that $20$ was about
% all we can handle before too many topics seemed too indistinct from
% each other to differentiate.

Our Topic Analysis tool is based around a third-party implementation
of LDA.  While we could have used LSI or similar techniques as the
engine, our preliminary studies found LDA to provide more promising
results for topic analysis.  We note that this step is a slow one, as
executing even one instance of LDA involves significant computation,
and we perform LDA once per window.  Figure \ref{fig:commits} shows
how LDA is applied to a set of messages, and how the topics are
extracted and related to the messages.

@@

Our input for topic similarity is the top 10 most common words of each
topic.  Each topic is compared to each other topic in the system,
given a certain threshold of top 10 word similarity, such as 8 out of
top 10 matching words.  10 words were chosen because often people care
about top 10 comparison and it allowed for some common words to exist
but not cause topics to match so closely.  We then apply the
transitive closure on similar topics to this topic similarity matrix;
this is similar to modelling topics as nodes and similarity as arcs,
then fill flooding along the similarity arcs until we have partitioned
the topics into clusters of similar topics. Figure \ref{fig:closure}
illustrates clustering of topics by topic similarity. These clusters
of topics are called \emph{trends}. 
Trends are interesting because they that topic during development has occurred over more than one period.

This approach does have a weakness, in that nodes that are a few neighbors
away in similarity might not share any similar words.  We use this
measure because we want to be able to color or highlight topics that
are related to each other and track them throughout time.

Once we have determined our similarity clusters we are ready to analyze
and to plot the topics.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.42\textwidth]{transitiveclosure}
  \caption{Topic similarity demonstrated by clustering topics by the transitive closure (connectedness)
    of topic similarity. Nodes are topics and
    arcs imply 
some notion of similarity, e.g.,  topics share 8 out of 10 top ten words.}
\label{fig:closure}
\end{figure}

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer
  Visualization is an integral part of our topics analysis which
  allows us to quickly explore the topics and trends of a project.
  Visualizing the results is desirable as it allows us to explore the
  data from different points of view to address different questions:
  How are the topics spread over time?  How many independent topics
  per period were there?  What repeating trends were there?  Did the
  trends dominate, or did local topics dominate?  How broken up were
  the trends?  
  Since we have a variety of questions to answer we suggest multiple
  visualizations, which we now describe.

%   Visualization provides us with a framework that allows us to
%   visually answer many questions about: the spread of topics, how many
%   topics were independent per period, what the repeating trends were,
%   if the trends dominated or did local topics dominate, how continuous
%   trends were, and what the topic text in topic, trend, or period was.
%   Since we have multiple questions to answer we have multiple
%   visualization, described within this section, that help answer
%   different questions.

We have devised several techniques for visualizing these topics and
trends.  For all of these techniques if we find trends that have
continuous segments, then we plot those segments as joined
horizontally across time. One technique is the \emph{compact-trend-view},
shown in Figure \ref{fig:topicsmear} and Figure \ref{fig:zoomedsmear},
that displays trends as they occur on the time-based x-axis
(placement along the y-axis is arbitrary).  Another technique is
the \emph{trend-time-line}, shown in Figure \ref{fig:trendtimeline}, where
each trend gets it own distinct y-value, while the x-axis is time;
these topics are then each plotted on their own line across time as
they occur. Our final technique is the \emph{trend-histogram}, shown in Figure
\ref{fig:histogram} where we plot each trend on its own line but stack
up the segments of the trend, much like a histogram. Each topic has its top 10 words 
listed in descending order of frequency from top to bottom, this text is embedded inside the topic's box.
A trend has all of its topic text embedded side by side within the same trend box, which at this resolution requires the reader to zoom in electronically.

 The \emph{compact-trend-view} (Figure \ref{fig:topicsmear})
  attempts to show all topics and trends at once across time in a
  compact view that could fit on one page.   The
\emph{compact-trend-view} (Figure \ref{fig:topicsmear}) tries to sink
the larger trends to the bottom.  Once the larger trends are stacked
we fill in the gaps with the smaller trends in the same window, and
then stack the smaller trends on top.  Although there is a chance that
gaps will be left due to non-optimal stacking, in practice there are
lot of small trends (90\% of all trends contain 1 topic) that fill in
these gaps quickly.  Different instances of the same trend share the
same color; apart from that, the color of trends is randomly chosen
and color similarity is not meaningful.  The compact-trend-view is
convenient because it shows all of the topic information, as shown by
the zoomed-in view of MaxDB 7.500's compact-trend-view in Figure
\ref{fig:zoomedsmear}. The \emph{compact-trend-view} makes repeating
continuous trends easy to pick out, although discontinuous trends are
harder to spot, and provides a general summary of the topics within a project.

@@

We applied our tools and visualizations to the repositories of three
open source database systems: PostgreSQL, MaxDB, and Firebird. The
total number of commits analyzed was over $66,000$.

@@

When we examined PostgreSQL's history from 1996 to 2004, which includes
over $20,000$ commits. We did not find many trends with two or more
topics, using a similarity of $7/10$. $7/10$ was chosen because it
preserve independent topics but seemed to be a threshold
value where more serious trends started to appear.

@@

We tracked Firebird from August 2000 to January 2006, we extracted comments from $38,000$ commits. We found that with a
similarity of $7/10$ Firebird had far more continuous and
recurring trends than PostgreSQL.  The first large trend was discontinuous
across time but explicitly references one author \texttt{carlosga05}
and words like \emph{added}, \emph{fixed}, and \emph{updated}.

@@

The query optimizer was a topic that recurred during MaxDB's
development. In our plots, topics that mention \emph{optimizer} occur
four times, yet in the global-trend-view (Figure
\ref{fig:statictopics}, explained in Section \ref{sec:compare}) it is not in any of the topics. A query
optimizer is an important component of an DBMS, but as we have shown
it does not appear as a topic on its own. We tried to remove words to
see if we could get an \emph{optimizer} topic. After removing stop
words and then two of the most common words, the global analysis
finally found a topic with optimizer in its topic ten words. Our
analysis shows that optimizer was important but it had been obscured
by the global topic analysis, which used the entire history of
messages, but would have been noticed using the more local topic
analysis such as our windowed topic analysis, which used a window of
messages.
